In this chapter, we identify a set of future directions for searches of dark matter at colliders. 
We note that this is not an exhaustive list, and gives only a personal perspective. 

%\subsection{Future evolution of models}

%What models are we developing? 
%- the scalar has low x-sec, will be more interesting when we can probe with mu < 1 
%- LLPs and why and how
%- t-channel are going away from the dominance of s-channel
%- SUSY electroweak is the future (Mike Hance's talk?)

In absence of new physics signals with large couplings with the LHC data so far, the obvious direction 
in terms of benchmark models for invisible particle searches at colliders is towards lower rate reactions
and signatures that are more difficult to detect. 

Extended scalar sectors are an example of such benchmark models. LHC Run-2 searches are only starting to be sensitive to the simplest (pseudo-)scalar models, and the development of further extensions is motivated by the expected sensitivity to this type of interactions. The HL-LHC dataset will bring a sensitivity of up to 3 TeV in mediator masses for unit couplings. 
%http://cms-results.web.cern.ch/cms-results/public-results/preliminary-results/FTR-16-005/CMS-PAS-FTR-16-005_Figure_008.png

Motivated by the wide variety of theoretically possible, experimentally challenging long-lived particle signatures, various efforts are ongoing to derive a prioritized set of benchmark models for collider searches in a similar fashion as Ref.~\cite{Abercrombie:2015wmb}. 
Among the various possibilities, we note the bottom-up approach adopted in~\cite{Buchmueller:2017uqu}, choosing masses and couplings for the models described in~\ref{sub:simplifiedModels} so that they must include a long-lived particle. 
The categorization of the models by production operator and final state permits to adopt a more systematic set of benchmarks for this kind of signatures, that can then be mapped onto more complete theories. 



for future searches for invisible particles is towards processes 

Models current under development 

Experimental holes. 

%\subsection{Future evolution of experimental searches}

The development of well-motivated models is crucial 

A generic event selection for excesses of \MET also provides an inclusive sample for more targeted searches as it will be discussed in Chapter~\ref{sec:05_Future}. 

Full-dataset HL-LHC results~\cite{Campana:2016cqm} will bring considerable improvements in the precision knowledge of the SM Higgs sector, with a sensitivity to Higgs to invisible decays from couplings measurements down to 10\%. %biggg dealllll  
The sensitivity to the electroweak production of SUSY particles will be increased, gaining from 50\% to 100\? in the superpartner mass reach. %(from 50 to 100\%)

The results of HL-LHC searches and measurements will be crucial to inform the construction of future colliders

\cite{Ilten:2016tkc} % LHCb turbo searches for dark boson
\cite{Ilten:2018crw} %Serendipity
\cite{Golling:2016gvc} % FCC-hh
 %HL-LHC AR
\cite{Liu:2017zdh} %FCC-ee, dark sector specifically

%\subsection{Future evolution of cosmological constraints}

Idea here: have a couple of sentences in future directions divided by chapter. One page max. 

%Higgs and Z portals
%For future: mention future e+e- colliders in VH production. 
%New: https://arxiv.org/abs/1712.07237
Mention inclusive vs exclusive searches, monojet is the basic event selection then one selects more things on top of that. 
%Link to:
%Moreover, the sensitivity hierarchy of \MET+X searches does not privilege the jet+\MET final state if there is a direct new physics coupling between a vector boson and the DM, as in the case of the EFT model mentioned in Section~\ref{sub:EFT}, or if the radiated object is a new particle~\cite{Autran:2015mfa}. 
%This latter signal motivates searches in the \MET+generic resonance final state CD: keep for later. 


However, no attempts have yet been made to connect the models in~\cite{Buchmueller:2017uqu} to cosmological history. 

Chapter 2



%too LLP? "we need this, an example of a way to do it is here"

Chapter 3

Remind people of how TLA was developed out of a hole, and in the same way develop simplified model connection to dark photons, are there holes?
- search for other scalars below the Higgs mass, lighter than the Z (1612.07282)
- this will need special triggering strategies, especially track triggers and triggerless readout (cite LHCb's MikeW)

Try and do searches with MET+ISR+weird stuff (or ISR is a new resonance like the Z' but do we really have to cite Whiteson?)

HL-LHC and future colliders a couple of sentences
- phil's stuff: under a particular set of assumptions, only the scalar will be left (i don't like that message)
- making use of highest lumi/CoM requires precise understanding of backgrounds
- some LLP (cite Curtin)

Chapter 4

Next generation of ID and DD, and maybe directional, will give us a lot more results on thermal wimps

%"We advocate that": not sure i want to put that in
A further step towards a more informative comparison between these results is to convey an idea of the impact of those uncertainties in the plots. 
The main uncertainties for LHC searches have been outlined in Sec.~\ref{03_ExperimentalResults}, and they are traditionally displayed as one-sigma bands around the LHC constraints. 
Additional uncertainties in the procedure that translate collider results in the DD and ID planes stem from the matrix elements of the nuclei, but this is an uncertainty that
affects DD as well. %Could make this clearer and cite? 
The situation is less straightforward when comparing collider results to measurements subject to astrophysical uncertainties (for a summary of DD and ID uncertainties, see~\cite{Feldstein:2014ufa,d300ef23986a49099715e661295a4d72} and references therein) as those are difficult to estimate and can have a large impact that affects different kinds of experiments differently. 
The experimental and theoretical communities have not yet agreed on a common presentation of experimental uncertainties, but we expect the discussion will continue in the future. 


It is also worth mentioning that, even though comparisons between collider, DD and ID are becoming the standard for these communities,
the comparison with LHC, DD and ID results with results from astrophysics beyond the relic density is a growing field. 
This is particularly important since all the observational evidence of DM that we have is gravitational,
so DM properties such as mass and density in our and other galaxies
can be inferred from galaxy simulations or deviations in astrophysical observables, 
and signatures of DM self-interactions in rotation galaxies or cluster collisions can complement and verify any DD 
observations. Even though we do not cover this topic in detail in this review we refer to~\cite{Buckley:2017ijx}. 



These are exciting times. %I kinda want to finish with this sentence