% template.tex, dated April 5 2013
% This is a template file for Annual Reviews 1 column Journals
%
% Compilation using ar-1col.cls' - version 1.0, Aptara Inc.
% (c) 2013 AR
%
% Steps to compile: latex latex latex
%
% For tracking purposes => this is v1.0 - Apr. 2013

\documentclass{ar-1col}
\usepackage[numbers]{natbib}

\setcounter{secnumdepth}{4}

% Metadata Information
\jname{Xxxx. Xxx. Xxx. Xxx.}
\jvol{AA}
\jyear{YYYY}
\doi{10.1146/((please add article doi))}


% Document starts
\begin{document}

% Page header
\markboth{Author et al.}{Short title}

% Title
\title{Dark Matter Searches at Colliders}


%Authors, affiliations address.
\author{Antonio Boveia,$^1$ Caterina Doglioni, $^2$
\affil{$^1$The Ohio State University + address}
\affil{$^2$Lund University + address}}

%Abstract
\begin{abstract}
Abstract text, approximately 150 words. 
\end{abstract}

%Keywords, etc.
\begin{keywords}
keywords, separated by comma, no full stop, lowercase
\end{keywords}
\maketitle

%Table of Contents
\tableofcontents


% Heading 1
\section{INTRODUCTION}
\label{sec:intro}

Define Standard Model (SM). 

%CD: need to decide whether Dark Matter or dark matter.

%\input{00_Examples}

%Heading 1
\section{REACTIONS FOR INVISIBLE PARTICLE SEARCHES AT THE LHC}

\input{02_Reactions}

%shared context: many possible new physics searches at the LHC
%problem: can't do them all
%solution: strong theoretical motivation, as well as observability
%exposition: particular case of DM

Collider experiments have a nearly unlimited number of theoretically motivated BSM search targets. 
Theoretical arguments are not sufficient for a new physics model to be tested at the LHC: 
couplings to SM particles need to feature in the model, to produce new particles and observe 
their signatures in the detectors. 

Dark matter is one of the priorities for the physics programs of LHC experiments
as a large number of its theoretical representations can be observed at the LHC detectors. 

%everyone thinks of WIMPs, how strong is strong, how weak is weak? quantitative question of coupling, depends on model. in the introduction: need to talk about DM properties. Weak enough that there is no visible EM signal (no light emission or absorption). Relate those properties to what the particle physics properties need to be. Have a model in mind: s-channel mediator between DM and SM, weakness of interaction comes from particle being heavy or coupling being small. DMF models have order=1 couplings. 

Models of particle dark matter generally include SM couplings to satisfy cosmological observations
in the freeze-out case. These couplings need to be weak enough that there is no visible signal of dark matter particles, as there is no evidence for dark matter interacting strongly with baryonic matter, 
nor for its emission or absorption of light. %i don't like its

The observations mentioned in Section~\ref{sec:intro} require the dark matter particle
to be stable on a cosmological timescale, so dark matter itself will be
invisible to colliders. However, the rest of the event is not: one can observe DM particles
due to their missing momentum in the transverse plane, if they recoil against one or 
more visible SM particles. 
%talk about LHC transverse momentum conservation

\clearpage


% Summary Points
\begin{summary}[SUMMARY POINTS]
\begin{enumerate}
\item Summary point 1. These should be full sentences.
\item Summary point 2. These should be full sentences.
\item Summary point 3. These should be full sentences.
\item Summary point 4. These should be full sentences.
\end{enumerate}
\end{summary}

% Future Issues
\begin{issues}[FUTURE ISSUES]
\begin{enumerate}
\item Future issue 1. These should be full sentences.
\item Future issue 2. These should be full sentences.
\item Future issue 3. These should be full sentences.
\item Future issue 4. These should be full sentences.
\end{enumerate}
\end{issues}

%Disclosure
\section*{DISCLOSURE STATEMENT}
If the authors have noting to disclose, the following statement will be used: The authors are not aware of any affiliations, memberships, funding, or financial holdings that
might be perceived as affecting the objectivity of this review. 

% Acknowledgements
\section*{ACKNOWLEDGMENTS}
Acknowledgements, general annotations, funding.

% References
%
% Margin notes within bibliography
\section*{LITERATURE\ CITED}

To download the appropriate bibliography style file, please see \url{http://www.annualreviews.org/page/authors/author-instructions/preparing/latex}. 

\\

\noindent
Please see the Style Guide document for instructions on preparing your Literature Cited.

The citations should be numbered in order of appearance. For example:





\begin{verbatim}
\begin{thebibliography}{96}
\expandafter\ifx\csname
natexlab\endcsname\relax\def\natexlab#1{#1}\fi

\bibitem{Glashow:1961tr}
Glashow SL. \textit{Nucl. Phys.} 22:579 (1961)

\bibitem{Weinberg:1967tq}
Weinberg S. \textit{Phys. Rev. Lett.} 19:1264 (1967)

\bibitem{Salam}
Salam A.  In \textit{Elementary Particle Theory: Relativistic
Groups and Analyticity. Proceedings of the 8th Nobel Symposium},
ed. N~Svartholm, p. 367. Stockholm: Almqvist \& Wiksell (1968)
\bibitem{Fritzsch:1973pi}
Fritzsch H, Gell-Mann M, Leutwyler H. \textit{Phys. Lett.} \textit{B} 47:365
(1973)

\bibitem{Gross:1973id}
Gross DJ, Wilczek F. \textit{Phys. Rev. Lett.} 30:1343 (1973)

\bibitem{Politzer:1973fx}
Politzer HD. \textit{Phys. Rev. Lett.} 30:1346 (1973)

\bibitem{Kronfeld:2010bx}
Kronfeld AS, Quigg C. \textit{Am. J. Phys.} 78:1081 (2010)

\bibitem{Langacker:2010}
Langacker P. \textit{The Standard Model and Beyond. Series in
High Energy Physics, Cosmology, and Gravitation}. Boca Raton:
CRC/Taylor \& Francis (2010)

\bibitem{GT2}
Quigg C. \textit{Gauge Theories of the Strong, Weak, and
Electromagnetic Interactions}. Princeton, NJ: Princeton
Univ. Press. 2nd ed. (2013)

\bibitem{MattS}
Schwartz MD. \textit{Quantum Field Theory and the Standard Model}.
Cambridge, UK: Cambridge Univ. Press (2013)

\bibitem{DynSM}
Donoghue JF, Golowich E, Holstein BR. \textit{Dynamics of the
Standard Model}. Cambridge, UK/New York: Cambridge
Univ. Press. 2nd ed. (2014)
\end{thebibliography}
\end{verbatim}

This coding results in the following formatted bibliography:


\begin{thebibliography}{96}
\expandafter\ifx\csname
natexlab\endcsname\relax\def\natexlab#1{#1}\fi

\bibitem{Glashow:1961tr}
Glashow SL. \textit{Nucl. Phys.} 22:579 (1961)

\bibitem{Weinberg:1967tq}
Weinberg S. \textit{Phys. Rev. Lett.} 19:1264 (1967)

\bibitem{Salam}
Salam A.  In \textit{Elementary Particle Theory: Relativistic
Groups and Analyticity. Proceedings of the 8th Nobel Symposium},
ed. N~Svartholm, p. 367. Stockholm: Almqvist \& Wiksell
 (1968)
\bibitem{Fritzsch:1973pi}
Fritzsch H, Gell-Mann M, Leutwyler H. \textit{Phys. Lett.} \textit{B} 47:365
(1973)

\bibitem{Gross:1973id}
Gross DJ, Wilczek F. \textit{Phys. Rev. Lett.} 30:1343 (1973)

\bibitem{Politzer:1973fx}
Politzer HD. \textit{Phys. Rev. Lett.} 30:1346 (1973)

\bibitem{Kronfeld:2010bx}
Kronfeld AS, Quigg C. \textit{Am. J. Phys.} 78:1081 (2010)

\bibitem{Langacker:2010}
Langacker P. \textit{The Standard Model and Beyond. Series in
High Energy Physics, Cosmology, and Gravitation}. Boca Raton:
CRC/Taylor \& Francis (2010)

\bibitem{GT2}
Quigg C. \textit{Gauge Theories of the Strong, Weak, and
Electromagnetic Interactions}. Princeton, NJ: Princeton
Univ. Press. 2nd ed. (2013)

\bibitem{MattS}
Schwartz MD. \textit{Quantum Field Theory and the Standard Model}.
Cambridge, UK: Cambridge Univ. Press (2013)

\bibitem{DynSM}
Donoghue JF, Golowich E, Holstein BR. \textit{Dynamics of the
Standard Model}. Cambridge, UK/New York: Cambridge
Univ. Press. 2nd ed. (2014)

\end{thebibliography}


\end{document}
