
%%Precision
Precision measurements of the Higgs boson properties and the comparison with SM theory also play a role in constraining the possible contributions to new physics, as 





This stuff tests Higgs portals.



$qq \rightarrow H qq$, $qq/gg \rightarrow VH$, $gg \rightarrow Hg$ Higgs production modes. 
%CD: This section is just CMS for now. Need to add ATLAS, but also cut as we're using too much space for this:
%https://atlas.web.cern.ch/Atlas/GROUPS/PHYSICS/PAPERS/HIGG-2015-03/fig_09.png
In all cases, in addition to a requirement of sizable missing transverse momentum, auxiliary visible objects are used to select the events. 
%Loop-induced signals are important. 
%For more info on importance and calculation of loops: 1605.08039, but we run out of citations and space
The events are divided in exclusive categories targeting specific production modes. The associated boson (VH) searches target the decays of Z bosons to electrons, muons light or heavy flavour quarks, while the W bosons can decay into light-flavour jets. 
The $qq \rightarrow H qq$ production mode is dominated by Vector Boson Fusion (VBF) processes, where the Higgs boson is produced in association with two hadronic jets that have a large pseudorapidity ($\eta$)
%CD: assume eta?
separation in the detector, and a large invariant mass. This topology is used to select events and discriminate between signal and background. 
%The large QCD backgrounds are suppressed by requiring that the missing transverse momentum recoils against the jets in the event. 
%If the missing transverse momentum was in the direction of the jets, there would be a chance of it coming from mismeasured jets. %CD: hmmm written this in a rush
The jet+MET search, described in more detail in the next section, is reinterpreted for the $gg \rightarrow Hg$ mode. 


can be probed through direct observation  

only direct measurements of the Higgs width are feasible, because [reasons]

%Pre-LHC constraints on the invisible Higgs width are derived from measurements of the ZH production channel at LEP in searches for new neutral Higgs-like bosons, where only the visible decays of the Z are observed. This is a common procedure to select events in LHC invisible particles searches. 
- how do we measure it now: LHC 
-- ofuck a hadron collider: how?
---- try to measure directly -> most recent results that are upper bounds, highlight the ZH mode
---- use precision measurements
-- how to connect: Dobrescu: make a bunch of assumptions because you don't have total width, measure coupling ratios, not model-independent anymore (but ratios make it less model-dependent)

In the SM, the Z boson can decay to a neutrino-antineutrino pair, while the Higgs boson decays into a pair of Z bosons each decaying to neutrinos. Additional decays of the Z and Higgs boson to particles beyond the SM modify the properties of the vector boson, such as width and couplings. 

%%MonoH

\textbf{Invisible decays of the H boson} within the SM only contribute to less than 0.1\% of the total decay width. For this reason, an observation of even a small contribution to the Higgs width from invisible particles would signal the presence of new physics phenomena that could be linked to invisible particles if 2\minvisible particles $< m_H$~\footnote{For the case of heavier invisible particles particles, see Ref.~\cite{Djouadi:2011aa}.}. 



%CD: need a number! 
It is not feasible to directly or indirectly measure the total and partial Higgs widths at a hadron collider and then extract the invisible contribution as done for the Z at LEP, as some of the decays (e.g. gluons and lighter quarks) have too large a background to be measured, the experimental resolution even for leptonic decays is large compared to the intrinsic Higgs SM width, and the kinematics of the ZH process is not fully determined as in lepton colliders. %CD: this is ambiguous also because I am not sure I fully understand the first and third points completely - need AB's help, page 2 of Dobrescu/Lykken. 
%Lykken/Dobrescu, 1210.3342: Total theoretical SM width/mass for H125: 3.2 * 10^-5 MeV, due to small Yukawa of b quark and suppression of WW*. From rates and couplings,  can extract upper and lower limits on the exotic Higgs branching fractions, which come from the upper/lower limit on the total width. This paper ignores exp uncertainties. 
%Wagner Dark Side of the Higgs boson: omit because we don't care about non-SM Higgses
%The width can be extracted from the lineshape in the low-background channel $Z \rightarrow ZZ \rightarrow 4l$, assuming a SM width. This is what CMS has done. 
%If one does not want to assume the SM width, one can still extract the width
%above 190 GeV where the experimental resolution is better. 
%what we want to see is a larger total width with less normalization because of the invisible decays


%The upper limit on the invisible BR from Higgs decays is 25%. 
%ATLAS Abstract
%Direct searches for invisible Higgs boson decays in the vector-boson fusion and associated production of a Higgs boson with W/Z (Z ? ??, W/Z ? jj) modes are statistically combined to set an upper limit on the Higgs boson invisible branching ratio of 0.25. The use of the measured visible decay rates in a more general coupling fit improves the upper limit to 0.23, constraining a Higgs portal model of invisible particles.
%%Precision
Precision measurements of the Higgs boson properties and the comparison with SM theory also play a role in constraining the possible contributions to new physics, as decays into invisible particles would reduce the SM Higgs production and decay coupling strengths~\cite{Khachatryan:2016vau,Englert:2011yb,Aad:2015pla}. 
%For the Higgs boson, the upper limit on the branching fraction to visible and/or invisible non-SM particles only using precision measurements is 34\%
%In case we want to say what limits these
%The main limitation for the measurement of the invisible width of the Higgs at the LHC is due to QCD uncertainties the Higgs production cross-section, which limits the sensitivity of these searches to roughly 10\% of the SM value. 




%CD: do we need to answer "what if not"? No one seems to care, but one could maybe think of using monojet off-shell (tiny tiny region) and precision constraints for the off-shell region too, a la dijet. Main point for the moment: DD covers this region so we don't have to. 















% leading to a measurement of the number of light neutrino families compatible with cosmology; if the partial widths of the decays into visible particles are subtracted from the total width, the invisible width can be measured to


%Searches for invisible particles at high-energy colliders are successful, since the Z boson branching fraction to light and weakly interacting particles is sizable. 
%The cross-section of Z to neutrinos at the LHC is [blah]. While included in the SM, these processes constitute a [testbed] and a background to further search for new invisible particles. 




%Higgs and Z portal models decay more in MET than they would do in the SM only



%%monojet

%Other benchmark scenarios such as compressed SUSY scenarios, 
%maybe explain?, 
%squark pair production, 
%who ordered that
%non-thermal singly-produced invisible particles, 
%and Large Extra Dimensions (ADD) are also constrained by the ATLAS and CMS searches, in some cases providing the most stringent constraints to date. 

%%What it means for the models we talked about
%Since no significant excess is found in any of the signal regions, limits are set on the parameter space of Higgs portal models described in Sec.~\ref{sec:HZPortalModels} and simplified models described in Sec.~\ref{sec:BSMMediatorModels}, namely where the SM-invisible particles interaction is mediated by $s-$channel vector (V), axial vector (AV), scalar (S) and pseudoscalar (P) and colored scalar mediators.  

%Sidebar (50 words minimum, 200 words maximum) briefly discussing a fascinating adjacent topic;
%insert below Literature Cited section, but indicate near which section in text the sidebar should be typeset

%\subsubsection{Searches with jets}
%monojet

%%CD: I am not sure I would want to read this summary of monojet search. But maybe I'm just jaded. Anyway, if we can, we should make it more interesting / give it a slightly different spin than just a plain description. 

%and  QCD subprocesses matter - too much detail too little space https://cds.cern.ch/record/159861/files/198507018.pdf


%%Sidebar (50 words minimum, 200 words maximum) briefly discussing a fascinating adjacent topic; insert below Literature Cited section, but indicate near which section in text the sidebar should be typeset
%\begin{textbox}[!h]
%\section{Precision estimation of background for \MET+X searches}
%In order to relate the number of events in the jet+\MET
%signal regions (where $Z\rightarrow \nu\nu$ dominates) and control
%regions (where events with jets produced in association with
%$Z\rightarrow ll$, $W\rightarrow l\nu$ and $\gamma$ are used to maximise
%the statistical power of the background estimation),
%one needs to rely on a precise theory  
%prediction of the ratio of the V+jets cross-sections. 
%This is why this is important, this is 10 words. 
%This is why this is important, this is 10 words. 
%This is why this is important, this is 10 words. 
%This is why this is important, this is 10 words. 
%This is why this is important, this is 10 words. 
%This is why this is important, this is 10 words. 
%This is why this is important, this is 10 words. 
%This is why this is important, this is 10 words. 
%This is why this is important, this is 10 words. 
%This is why this is important, this is 10 words. 
%%\subsubsection{Precise background estimation}
%%\label{sub:precision}
%
%%Cite AR on this topi (but it's 2009): \cite{doi:10.1146/annurev.nucl.56.100704.122617}
%%from ooutline
%%- LHC results				
%%	citations for most recent
%%	- Differences between ATLAS and CMS:				
%%		- CMS starts with only Z, ATLAS uses Z and W, CMS uses everything including gamma				
%%		- theoretical issues with using Z and W				
%%		- Pozzorini paper: shit is complicated, W and Z are one thing but if you want to do photon it's a different story				
%%			- dependence of result on analysis cuts				
%%			- QCD correction				
%%			- EW corrections				
%\end{textbox}

%in OOutline we wanted to quote example numbers, but there is a lot of eyeballing
%>1000 GeV: EM10: 226+/-16 events predicted, 245 observed
%WIMP minvisible particles 400, mmed 1000: 0.2 (eyeballed)*200 GeV  



%CD: Maybe we put this in the reactions chapter? It is quite an important statement


%monoH come from CMS but maybe ATLAS has a better reinterpretation. 
The most stringent 95\% C.L. observed (expected) upper limits on the invisible branching
fraction from jet+\MET searches are 53\% (40\%).  TODO look for ATLAS?
%(CMS, combining jet and vector boson radiation categories). 
%V/AV come from CMS search, ATLAS is less sensitive as it's 1.55 TeV
Vector and axial vector mediators are excluded by LHC searches at values of \minvisible particles up to 700 and 400 GeV respectively with \mmed up to 1.8 TeV. This choice of model and couplings produces a relic density that is lower than the Planck measurement and it is still unconstrained by LHC searches for \minvisible particles$>$0.3 TeV at \minvisible particles$=$1.8 TeV for the vector mediator, and for 0.65$<$\minvisible particles$<$0.75 TeV at \minvisible particles=1.8 TeV for the axial vector mediator\footnote{Here and in the following, we quote observed limits at 95\% C.L. and refer to the bibliography for expected limits and 90\% C.L. limits.}. 
%CD: Not sure this is interesting for anyone? A bit complicated to project a 2D plot in words
%pseudoscalar comes from CMS
The LHC limit on the pseudoscalar mediator mass is lower due to the Yukawa-like couplings suppressing the cross-section with respect to spin-1 mediators, and it is 0.4 TeV in the CMS search for \minvisible particles up to roughly 150 GeV. 
Jet+\MET searches are not yet sensitive to scalar mediators with the chosen couplings. 
%t-channel comes from ATLAS
%CMS
%Colored scalar mediators with masses up to 1.4 TeV at values of \minvisible particles = 60 GeV are excluded.
%ATLAS
%CD TODO: check what parameters are people using?
Colored scalar mediators with masses up to 1.7 TeV at values of \minvisible particles = 10 GeV are excluded for \ginvisible particlesq=1 and \minvisible particles=100 GeV. Considering this exclusion limit, this model still provides a viable invisible particles relic density for \mmed \minvisible particles above roughly 500 GeV at \mmed=1.7 TeV\footnote{The ATLAS and CMS results do not use the same parameters, here we report the ATLAS result.}.




This observable is corrected for detector effects

Moreover, the ATLAS Collaboration has used the ratio of cross sections of events containing a jet and \MET and events containing a jet produced in association with an opposite-sign same-flavour dilepton pair from the decay of a Z/$\gamma*$ boson~\cite{Aaboud:2017buf}, corrected for detector effects. 
%in a fiducial phase space - saving space, leaving it unsaid
This is an observable sensitive to the anomalous production of events with jets and \MET, and uses many of the techniques from the the jet+\MET search described above to estimate background. The constraints derived are comparable to those of the jet+\MET search with the equivalent dataset. Unlike most other searches for new physics described in this review, detector effects are already accounted for (\textit{unfolded}) when presenting results, so that there is no need to implement a detector simulation to reinterpret this search. 


\subsection{monophoton/V}



differences:
+ Higgs to invisible (not really because monojet also)
+ the EFTs 

%i'm ok with Zll just being cited earlier, kayla will be sad but 

%not sure we need this marginnote
%\begin{marginnote}[]
%\entry{Jet substructure}{techniques employed to extract information from the radiation %pattern of a jet, by analyzing its constituents or its reconstruction history. For a review, see ~\cite{Larkoski:2017jix}. 
%}
%\end{marginnote}

% needs transition now what? backkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkk-edkkkkiskkinyourofficenownogoodathomerightiforgot

ok let's cut this shit


%Don't want Zgamma resonances so restrict mllgamma < 1 TeV, but maybe too much info
For the \textbf{photon+\MET search} at ATLAS and CMS~\cite{Aaboud:2017dor,CMS-PAS-EXO-16-014}, jets and leptons faking photons are estimated directly from data.
The $\gamma$+jet background where the jet is mismeasured and produced \MET is suppressed by the requirement that the photon and the direction of the \MET vector do not overlap in the azimuthal plane. 
The total uncertainty is dominated by the statistical uncertainty in the control regions, ranging from 4\% to 10\%. 
%Total ATLAS uncertainty in SR1:
%post-fit
%>>> 160./2600
%0.06153846153846154
%pre-fit
%>>> 200./2400
%0.08333333333333333

%Detail detail detail
%The \textbf{photon+\MET searches} use a photon to trigger the events to be recorded for analysis, and selects events containing an isolated photon above 150 GeV and no leptons. The number of events from Z decays to neutrinos in association with a photon can be estimated in events where the lepton veto is inverted and the contribution of visible Z and W boson decays is removed from the transverse momentum balance of the event, and transferred to the signal region. 

ATLAS and CMS \textbf{W/Z+\MET searches} where the vector boson (V) decays to a quark-antiquark pair specifically select events where the decay products from the high-\pt{} boson are collimated, to better discriminate signal and background. 
QCD jets will not present any \textit{substructure}, while the decay products of vector bosons grouped into large-radius jets have a typical two-prong pattern from the hadronization of the quark-antiquark pair. 
The dominant background is still Z decays to neutrinos in association with jets, followed by W decays where the lepton is not identified, and top quark decays. 
The shapes of these backgrounds are estimated using simulation, while the normalization is determined in control regions, similarly to the jet+\MET search. 
%In the CMS search, the V+\MET backgrounds are estimated in a simultaneous fit together with the jet+\MET backgrounds. 
The main uncertainty for this search (up to 9\%~\cite{Sirunyan:2017jix} and 13\%~\cite{Aaboud:2016qgg}) %ATLAS, CMS is 9\% due to the tagging requirements
is due to the modeling of the substructure observables. 

\begin{marginnote}[]
\entry{Jet substructure}{a set of techniques employed to extract information from the radiation pattern of a jet, by analyzing its constituents or its reconstruction history. For a review, see ~\cite{Larkoski:2017jix}.}
\end{marginnote}

The \textbf{Z+\MET searches} where the Z boson decays leptonically are sufficiently general to be sensitive to simplified models of invisible particles with a Z radiation, as well as to Higgs decaying into new invisible particles and produced in association with a Z~\cite{Sirunyan:2017qfc, Aaboud:2017bja}. 
The event selection includes a constraint on the dilepton invariant mass, which limits the backgrounds to diboson, leptonically decaying top quarks, Drell-Yan production and a small amount of triboson processes. 
The estimation of the main $ZZ\rightarrow 2\nu 2l$ background (about 60\% of the total backgrounds) uses simulation, as the data sample that could be used to constrain the normalization as in the jet+\MET search is statistically-limited. The main uncertainty for this search (10\% on the background estimation) comes from the theoretical uncertainties on this background. 
%who gives a shit fuck machine learning really
%The CMS search uses a Boosted Decision Tree (BDT) applied to events with the missing transverse momentum between 100 and 130 GeV, to enhance the sensitivity to invisible Higgs decays. 

%~\footnote{Leptonic decays W bosons have also been employed in the past for this kind of searches, but due to the additional experimental challenges (e.g. the presence of an additional invisible particle, the neutrino in the W decay) and the reduced sensitivity with respect to the hadronic decays, they have not been specifically pursued as invisible particles searches for the LHC Run-2.}

%ATLAS and CMS have pursued searches for missing transverse momentum produced in association with a photon%monophoton - if we're running low on citation space, remove CMS monophoton
%\cite{Aaboud:2017dor,CMS-PAS-EXO-16-014},%less lumi, unpublished, ,
%vector boson decaying hadronically %monoV, had, 2015+2016 (CMS) and 2015 (ATLAS)
%\cite{Sirunyan:2017jix,Aaboud:2016qgg} or leptonically %monoV, lep, 2015+2016
%\cite{Aaboud:2017bja, Sirunyan:2017qfc}. 

%One of the advantages of these search signatures over the jet+\MET one
%is the lower event selection threshold, thanks to the additional handles to suppress background
%provided by either the photon ISR or the leptonic decays. 

%Searches for invisible particles produced in association with a jet are the most sensitive among the searches employing an object radiated in the initial state, due to the large signal rates from the radiation of a gluon as opposed to the radiation of a photon or a W/Z/Higgs boson. The jet+\MET final state is however also affected by the largest SM and instrumental background, and only covers signals producing a high \MET to comply with data-taking limitations at the trigger level, due to the high-rate backgrounds producing signal-like signatures. It is therefore worth considering other objects as ISR, as those searches will be subject to different backgrounds, different kinds of systematic uncertainties, lower \MET thresholds, and can provide confirmation in case of an excess in the jet+\MET final state~\cite{Birkedal:2004xn,Petriello:2008pu}. 
%Gershtein:2008bf 2nd monophoton paper, save cites
%The sensitivity hierarchy of \MET+X searches does not necessarily privilege the jet+\MET final state if there is a direct new physics coupling between a vector boson and the invisible particles, as in the case of the EFT model mentioned in Section~\ref{sub:EFT}, or if the radiated object is a new particle~\cite{Autran:2015mfa}. %mono-Z-prime
%This latter signal motivates searches in the \MET+generic resonance final state CD: keep for later. 

[transition? or we've said that before?]

In absence of any excess over the estimated background, the photon+\MET searches provide the next-to-most stringent constraints on invisible particle production, with 95\% CL cross-section limit constrained to be below 2.5 and 7~\fb depending on the \MET threshold~\cite{Aaboud:2017dor} ranging from 150 to 300 GeV. 

%we need to qualify because otherwise it looks like it's better than monojet, but then isn't it?
These results are interpreted in the context of EFT models where the invisible particles couple directly to the photon~\cite{Petrov:2013nia,Berlin:2014cfa}, excluding EFT scales between 150 and 750 GeV for \mdm=100 GeV, assuming the maximal coupling value allowed by perturbativity. %is this right? the "direct" vertex is a consequence of the EFT after all. should we give the coupling value?

axial-)vector simplified models for invisible particles,
The rates of 

Mediators 







JUNK: Run-1 LHC searches privileged EFT operators, while Run-2 searches are limited to using EFT models with a SM singlet and a boson pair, coupled to invisible particles through a contact interaction (see e.g.~\cite{Petrov:2013nia,Berlin:2014cfa}). These models provide a direct SM-invisible particles interaction and motivates searches in final states with a vector boson accompanied by \MET. 


in the region where the mediator can decay to invisible particles. %Not mentioning off-shell, otherwise it's a closed region. 
The ATLAS photon+\MET search also sets limit on the EFT model where the invisible particles couples directly to the photon, 
excluding EFT scales between 150 and 750 GeV for \minvisible particles=100 GeV, assuming the maximal coupling value allowed by perturbativity. 
%The excluded region decreases to 150-600 GeV for a coupling value of 3. 
Searches in the \MET+hadronic Z final state constrain vector and axial vector mediator
masses of \mmed $<$ 650 GeV for \minvisible particles=100 GeV\footnote{As a side note that is useful to compare results from different LHC datasets, 
Run-1 V+\MET searches used a version of the vector simplified model
that enhanced W radiation because of the constructive interference 
due to different up- and down-quark mediator couplings,
but that was not gauge invariant~\cite{Bell:2015sza,1475-7516-2016-01-051}.}. 
Searches in the \MET+leptonic Z final state provide constraints on 
\mmed $<$ 650 GeV for \minvisible particles=100 GeV for vector and axial vector mediators. 
W/Z+\MET searches are also uniquely sensitive to the radiation of a boson 
from the mediator particle in the case of colored scalar models~\cite{Bell:2012rg}, 
but the Run-2 searches do not present this interpretation. 
%CD: maybe we add a link to the 8 TeV search but it's worse than monojet 


%- however other models. eg.. eFT

%In absence of any excess over the estimated background, 

%Other things we aren't mentioning:
%- Bell's "ISR from mediator" (but no one checks that; this is the 'from the dark interaction') 
%- the monoW fake strong limits from Run-1 but they were wrong anyway
- (for before -> going there and adding)
-- UA1 monojets, can stick it in "difficulties of precision"
---- question here: the paper to cite is here:
https://cds.cern.ch/record/159861/files/198507018.pdf
and basically says "UA1 has no fucking clue about backgrounds in general" 
QUESTION HERE FLASH FLASH BLINK BLINK -> do we want to use citations to make fun of Rubbia?
read this: https://books.google.se/books?id=iguOgUlYMHoC&pg=PA22&lpg=PA22&dq=UA1+monojet&source=bl&ots=gbHPTHKsi7&sig=dI4slozaz4Wj1LBqDWliEVb0Ps0&hl=en&sa=X&ved=0ahUKEwiUl56PmY3ZAhXJh6YKHUiSB3YQ6AEINzAC#v=onepage&q=UA1%20monojet&f=false
ok so we use this to introduce the important of precise predictions
ok
not now though, i put a pointer 

-- t-channel results
maybe we should put this in (fuck that whitepaper i have no bandwidth)
yes
we should
but in: experimental challenges: how hard it is when you have this cocktail of stuff 
you mean experimental? it seems like SUSY does fine. the MC generation i'd say too hard to explain
however, though this wa first a problem in monojets, it is now a standard problem and the reason why we all use MC
THIS IS A 30 year old problem that is solved now
define this?
the ellis paperaha
i thought you were talking tchannel
ok what about tchannel

https://atlas.web.cern.ch/Atlas/GROUPS/PHYSICS/PAPERS/EXOT-2016-27/fig_08.png
"this analysis sets limits on colored scalar mediators"
tbh they are the Worlds Best Records


%%MonoHiggs

%CD: we already said that before
The newly discovered Higgs boson is of particular interest for invisible particles searches at the LHC. 
Higgs radiation is kinematically and PDF suppressed, but searches for a mono-Higgs have
other strong theoretical motivations. 
Higgs portal models are the simplest incarnation of theories where the coupling
between the dark sector and the SM is realized through a Higgs boson. Higgs couplings
to at least a new scalar are necessary for the gauge invariance of simplified models described in
the previous chapters towards more complete theories. 
%The rates of signatures including Higgs bosons are small, but 
Gauge symmetries link Higgs+\MET signatures and signatures including W, Z bosons
or jets as well as two-body mediator searches~\cite{Liew:2016oon}. 
%so results from Higgs searches are not to be taken in isolation.  


derail somewhere then do monob then do SUSY which is pretty much done by Frederik and we keep LLP for tomorrow?
question 
mark
i still want to be awake until 5' of superbowl because i am curious about your commentary
'''';;;;; ok
i need: coffee + a walk or something
walking is good
can you walk to coffee 
 could go to the lake
 then tell me if the resident heron is there

it is america: no walking to coffee unless the city
it is snowing and icy and an alert about the weather buzzed my phone
hence i did not go in today
so maybe no driving anywhere to walk
also i function for another couple of hours 
so you have say 20' to go to the lake
take pictures?
i don't know if it's iced
iced latke


not enough time
coffee alone takes 10'
how do you make coffee
it is probably a complicated procedure
involving lambics
no that is a beer
alambicchi
https://naturaecultura.files.wordpress.com/2015/01/scansione0003.jpg
your coffee apparatus
A Large Alambic Coffee Apparatus (ALACA)

so i will walk t
o throw out trash
careful about ladies in cars throwing out trash and waitressing or trashing or running over 
probably waitressing is the least dangerous
tell me if lake is icy
the fish are still under there but the heron can't reach them

do not endanger yourselfs
good
do we want to get raaackedn
letsgetletsgetletsgetletsgetracked